---
title: "Presentation of GreenEdge data"
subtitle: "An overview"  
author: "Philippe Massicotte et al."
date: "2019-03-22"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
duo(
  code_font_family = "Fira Code",
  code_font_url    = "https://cdn.rawgit.com/tonsky/FiraCode/1.204/distr/fira_code.css"
)
```

# GreenEdge data

--

- Many sources of data:

--

  - Different formats (xls, csv, mat, bin, etc.)

--

  - Acquired at different temporal/spatial scales

--

  - Data structure highly variable among the researchers

--

  - **Difficult to exploit the data** (merge, visualization, etc.)

<center>
<img src = "https://media.giphy.com/media/l1KVaj5UcbHwrBMqI/giphy.gif">
</center>

---

# Importance of well-structured data

> Most data scientists spend only 20% of their time on actual data analysis and 80% of their time finding, cleaning, and reorganizing huge amounts of data, which is an inefficient data strategy.

<small>
https://www.infoworld.com/article/3228245/the-80-20-data-science-dilemma.html
</small>

--

We put a lot of effort into standardizing the data collected during the GreenEdge campaigns, so you do not have to.

Most of the job has been done by these persons:

1. **Marie-Pier Amyot**

2. Joannie Ferland

3. Etienne Ouellet

4. Flavienne Bruyant

---

# What has been done

- Screening the data for possible errors

  - Outliers, duplicated information, etc.
  
- Regrouping similar data into single files

  - Ex.: nutrients, absorption
  
---

# Data formatting

Some rules were used to format the data:

1. Only use small case letters in variable names.

2. No special characters (#$%?&*éê...)

3. Dates should be ISO format: yyy-mm-dd HH:MM:SS.

4. Units should be added at the end of the variable name (Ex.: `depth_m`).

5. Data should be tidy (i.e.: lines = observations, columns = variables).

---

# Data formatting

Each dataset contains a minimum number of columns that can be used to merge the data. **For the ice camp**:

1. date

2. date_time

3. latitude (`67.47973333`)

4. longitude (`-63.78953333`)

5. sample_type (`water`, `ice`, `meltpond`, ...)

6. sample_source(`niskin` or `underice`, ...)

7. depth_m (NA if sample_type is ice)

8. snow_thickness (`thin_snow` or `thick_snow`)

9. mission (`ice_camp_2015`, `ice_camp_2016`)

10. PI

---

# Data papers

**Two data papers** in preparation:

- Ice camp 2015-2016: Joannie Ferland

- Amundsen 2016: Flavienne Bruyant

- For possible publication in **Earth System Science Data (ESSD)** (I.F.: 8.8)

> Earth System Science Data (ESSD) is an international, interdisciplinary journal for the publication of articles on original research data (sets), furthering the reuse of high-quality data of benefit to Earth system sciences.

- Formatted data will be deposed in a repository (possibly LEFE-CYBER) **along with the original data**.

- Data will be citable using the DOI's of the data papers.

---

# Questions or comments?

<center>
<img src = "https://media.giphy.com/media/86nJuFHH25Uu4/giphy.gif">
</center>